---
title: "Caracal_Sightings_GIS"
author: "Jana Kemp"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse) #For general data handling
library(sf) #For spatial data handling
library(leaflet) #For interactive mapping
library(magrittr)   # Pipe operator (%>%) for clean code
library(htmltools) #For labels and popups
library(dplyr)
```

# Urban Edge Polygon
Reading in Cape Peninsula Urban Edge shape file and transforming it to Coordinate reference system 4326 (Uses longitude and latitude)

```{r}
edge <- st_read("C:/Users/27761/Documents/GIT/Caracal_GIS_Assignment/urban_edge/urban_edge/urban_diss.shp")

# Transform to WGS 84 (EPSG: 4326)
edge_sf <- st_transform(edge, crs = 4326)
```

# Reading and Transforming Caracal Data
Import a CSV file of caracal sightings reported to the Urban Caracal Project (Anonymised).
 

```{r}
caracal_data <- read.csv("C:/Users/27761/Documents/GIT/Caracal_GIS_Assignment/Caracal_Data.csv")

#Fixing error in dataset. Longitude and Latitude were mislabelled

caracal_data <- caracal_data %>%
  rename(longitude = Latitude)

caracal_data <- caracal_data %>%
  rename(latitude = Longitude)


#Editing mistake in longitude (was negative)
caracal_data[159, "longitude"] <- 29.75148

#Removing sightings far East of the Cape Peninsula (Not relevant)

caracal_data <- caracal_data[caracal_data$longitude <= 20, ]

caracal_sf <- st_as_sf(caracal_data, coords = c("longitude", "latitude"), crs = 4326)
```

# Creating an interactive map
Create an interactive map centered around Cape Town, added the urban edge as a polygon layer, and plot caracal sightings as orange circles. 
```{r}
leaflet() %>%
  addTiles(group = "Default") %>%
  #Setting the view to the Cape Peninsula
  setView(lng = 18.4, lat = -34.25, zoom = 9) %>% 
  #adding the urban edge polygon
  addPolygons(data = edge, 
              color = "black",       
              weight = 2,           
              fillColor = "grey", 
              fillOpacity = 0.5)  %>%  
  #adding caracal sightings
  addCircleMarkers(data = caracal_sf,
                   group = "Caracal",
                   radius = 5, 
                   color = "orange",
                   #labeling by date 
                   label = ~ as.character(caracal_sf$Date)) 
```

